---
title: "Overall_DA_Analysis_Revisions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(ggbeeswarm)
library(gridExtra)
library(cowplot)
library(kableExtra)
library(DT)
library(reshape2)
library(pheatmap)
library(RColorBrewer)

remove_rare_features <- function( table , cutoff_pro=0.1, parallel=1 ) {
  if(cutoff_pro<=0){
    message("No filtering will be done due to cutoff_pro set to 0 or below")
    return(table)
  }
  row2keep <- c()
  cutoff <- ceiling( cutoff_pro * ncol(table) )
  if(parallel <= 1){
    for ( i in 1:nrow(table) ) {
      row_nonzero <- length( which( table[ i , ]  > 0 ) )
      if ( row_nonzero > cutoff ) {
        row2keep <- c( row2keep , i)
      }
    }
    return( table [ row2keep , , drop=F ])
  }else{
    `%dopar%` <- foreach::`%dopar%`
    cl <- parallel::makeCluster(parallel)
    doParallel::registerDoParallel(cl)
    message("Running with: ", foreach::getDoParWorkers(), " cores")

    res <- foreach::foreach(i = 1:nrow(table), .combine = c) %dopar% {
      row_nonzero <- length( which ( table[ i , ] > 0))
      if(row_nonzero > cutoff){
        i
      }else
        NULL
    }
    parallel::stopCluster(cl)
    return( table[res, , drop=F])

  }

}
```

# Load in DA results

## Breast

```{r}
Breast_ASV_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/ASV_ALDEX2_age.RDS")
Breast_ASV_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/ASV_ANCOM_age.RDS")
Breast_ASV_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/ASV_Corncob_age.RDS")
Breast_ASV_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/ASV_MAASLIN2_age.RDS")

Breast_GENUS_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/Genus_ALDEX2_age.RDS")
Breast_GENUS_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/Genus_ANCOM_age.RDS")
Breast_GENUS_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/Genus_Corncob_age.RDS")
Breast_GENUS_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Breast/Genus_MAASLIN2_age.RDS")
```

## Prostate

```{r}
Prostate_ASV_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/ASV_ALDEX2_age.RDS")
Prostate_ASV_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/ASV_ANCOM_age.RDS")
Prostate_ASV_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/ASV_Corncob_age.RDS")
Prostate_ASV_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/ASV_MAASLIN2_age.RDS")

Prostate_GENUS_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/Genus_ALDEX2_age.RDS")
Prostate_GENUS_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/Genus_ANCOM_age.RDS")
Prostate_GENUS_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/Genus_Corncob_age.RDS")
Prostate_GENUS_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/Prostate/Genus_MAASLIN2_age.RDS")

#filter out structural zeros due to low sample size
Prostate_ASV_ancom$out <- Prostate_ASV_ancom$out[-which(is.infinite(Prostate_ASV_ancom$out$W)),]
```

## Colon

```{r}
Colon_ASV_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/ASV_ALDEX2_age.RDS")
Colon_ASV_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/ASV_ANCOM_age.RDS")
Colon_ASV_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/ASV_Corncob_age.RDS")
Colon_ASV_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/ASV_MAASLIN2_age.RDS")

Colon_GENUS_aldex <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/Genus_ALDEX2_age.RDS")
Colon_GENUS_ancom <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/Genus_ANCOM_age.RDS")
Colon_GENUS_corncob <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/Genus_Corncob_age.RDS")
Colon_GENUS_maaslin2 <- readRDS("~/Private/CHAPTER_4/Revised_data/DA/PATH_Retro/CRC/Genus_MAASLIN2_age.RDS")

#filter out structural zeros due to low sample size
Colon_ASV_ancom$out <- Colon_ASV_ancom$out[-which(is.infinite(Colon_ASV_ancom$out$W)),]
Colon_GENUS_ancom$out <- Colon_GENUS_ancom$out[-which(is.infinite(Colon_GENUS_ancom$out$W)),]
```

# Load in mettadata and taxa abundances

```{r}

#Load in Metadata
Combined_data <- readRDS("~/Private/CHAPTER_4/Revised_data/Combined_data.RDS")
Combined_data_filt <- readRDS("~/Private/CHAPTER_4/Revised_data/Combined_data_filt.RDS")


Combined_data <- Combined_data %>% select(-A_SDC_AGE_CALC, -PM_STANDING_HEIGHT_AVG, -PM_WAIST_HIP_RATIO,
                                          -REFINED_GRAIN_SERVINGS_DAY_QTY, -A_SDC_GENDER,
                                          -NUT_VEG_DAY_QTY)

Combined_data_filt <- Combined_data_filt %>% select(-A_SDC_AGE_CALC, -PM_STANDING_HEIGHT_AVG, -PM_WAIST_HIP_RATIO,
                                          -REFINED_GRAIN_SERVINGS_DAY_QTY, -A_SDC_GENDER,
                                          -NUT_VEG_DAY_QTY)


#Load in ASVs
ASV_table <- read.table("~/Private/Previous_Cancer_Project/RERUN_REV1/deblur_output/asv_table/feature-table.tsv",
                        sep="\t", quote="", comment.char="", row.names=1, header=T, skip=1, check.names = F)
colnames(ASV_table)

#Load in Genus table
Genus_table <- read.table("~/Private/Previous_Cancer_Project/RERUN_REV1/taxa/level6/feature-table.tsv",
                          sep="\t", quote="", comment.char="", row.names=1, header=T, skip=1, check.names = F)

#divide data up into different types

Breast_data <- Combined_data %>% filter(Type == "Breast")
Prostate_data <- Combined_data %>% filter(Type == "Prostate")
CRC_data <- Combined_data %>% filter(Type == "Colon")

#divide out asv and genus tables
Breast_ASV <- ASV_table[,Breast_data$X]
Prostate_ASV <- ASV_table[,Prostate_data$X]
Colon_ASV <- ASV_table[,CRC_data$X]

Breast_Genus <- Genus_table[,Breast_data$X]
Prostate_Genus <- Genus_table[,Prostate_data$X]
Colon_Genus <- Genus_table[,CRC_data$X]
```

# Generate CLR and RA tables

## CLR

Generated with a pseudocount of 1

### ASV

```{r}
ASV_Breast_CLR <- data.frame(apply(Breast_ASV + 1, 2, function(x){log(x) - mean(log(x))}), check.names = F)
ASV_CRC_CLR <- data.frame(apply(Colon_ASV + 1, 2, function(x){log(x) - mean(log(x))}), check.names = F)
ASV_Prostate_CLR <- data.frame(apply(Prostate_ASV + 1, 2, function(x){log(x) - mean (log(x))}), check.names = F)

#flip them for plotting
ASV_Breast_CLR_flip <- data.frame(t(ASV_Breast_CLR), check.names = F)
ASV_Breast_CLR_flip$X <- rownames(ASV_Breast_CLR_flip)

ASV_CRC_CLR_flip <- data.frame(t(ASV_CRC_CLR), check.names = F)
ASV_CRC_CLR_flip$X <- rownames(ASV_CRC_CLR_flip)

ASV_Prostate_CLR_flip <- data.frame(t(ASV_Prostate_CLR), check.names = F)
ASV_Prostate_CLR_flip$X <- rownames(ASV_Prostate_CLR_flip)
```

### Genus

```{r}
Genus_Breast_CLR <- data.frame(apply(Breast_Genus + 1, 2, function(x){log(x) - mean(log(x))}), check.names = F)
Genus_CRC_CLR <- data.frame(apply(Colon_Genus + 1, 2, function(x){log(x) - mean(log(x))}), check.names = F)
Genus_Prostate_CLR <- data.frame(apply(Prostate_Genus + 1, 2, function(x){log(x) - mean(log(x))}), check.names = F)

Genus_Breast_CLR_flip <- data.frame(t(Genus_Breast_CLR), check.names = F)
Genus_Breast_CLR_flip$X <- rownames(Genus_Breast_CLR_flip)

Genus_CRC_CLR_flip <- data.frame(t(Genus_CRC_CLR), check.names = F)
Genus_CRC_CLR_flip$X <- rownames(Genus_CRC_CLR_flip)

Genus_Prostate_CLR_flip <- data.frame(t(Genus_Prostate_CLR), check.names = F)
Genus_Prostate_CLR_flip$X <- rownames(Genus_Prostate_CLR_flip)
```

## RA

### ASV
```{r}
ASV_Breast_RA <- sweep(Breast_ASV, 2, colSums(Breast_ASV), '/')
ASV_CRC_RA <- sweep(Colon_ASV, 2, colSums(Colon_ASV), '/')
ASV_Prostate_RA <- sweep(Prostate_ASV, 2, colSums(Prostate_ASV), '/')

ASV_Breast_RA_flip <- data.frame(t(ASV_Breast_RA), check.names=F)
ASV_Breast_RA_flip$X <- rownames(ASV_Breast_RA_flip)

ASV_CRC_RA_flip <- data.frame(t(ASV_CRC_RA), check.names = F)
ASV_CRC_RA_flip$X <- rownames(ASV_CRC_RA_flip)

ASV_Prostate_RA_flip <- data.frame(t(ASV_Prostate_RA), check.names = F)
ASV_Prostate_RA_flip$X <- rownames(ASV_Prostate_RA_flip)
```

### Genus

```{r}
Genus_Breast_RA <- sweep(Breast_Genus, 2, colSums(Breast_Genus), '/')
Genus_CRC_RA <- sweep(Colon_Genus, 2, colSums(Colon_Genus), '/')
Genus_Prostate_RA <- sweep(Prostate_Genus, 2, colSums(Prostate_Genus), '/')

Genus_Breast_RA_flip <- data.frame(t(Genus_Breast_RA), check.names = F)
Genus_Breast_RA_flip$X <- rownames(Genus_Breast_RA_flip)

Genus_CRC_RA_flip <- data.frame(t(Genus_CRC_RA), check.names = F)
Genus_CRC_RA_flip$X <- rownames(Genus_CRC_RA_flip)

Genus_Prostate_RA_flip <- data.frame(t(Genus_Prostate_RA), check.names = F)
Genus_Prostate_RA_flip$X <- rownames(Genus_Prostate_RA_flip)
```


### Get list of non-rare taxa

Use prevelance cutoff of 5%
```{r}
Genus_Breast_filt <- remove_rare_features(Breast_Genus, cutoff_pro = 0.05)
Genus_Prostate_filt <- remove_rare_features(Prostate_Genus, cutoff_pro = 0.05)
Genus_CRC_filt <- remove_rare_features(Colon_Genus, cutoff_pro = 0.05)

ASV_Breast_filt <- remove_rare_features(Breast_ASV, cutoff_pro = 0.05)
ASV_Prostate_filt <- remove_rare_features(Prostate_ASV, cutoff_pro = 0.05)
ASV_CRC_filt <- remove_rare_features(Colon_ASV, cutoff_pro = 0.05)
```



# Heatmap functions

## Data Prep
```{r}

Comb_p_vals <- function(aldex,ancom,corncob,maaslin2, feat_list){
  aldex$feature <- rownames(aldex)
  aldex <- aldex[intersect(feat_list,rownames(aldex)),]
  ## we need to sort them all into the same order
  if(sum(grepl("ControlCase", colnames(aldex)))){
    aldex_sub <- aldex[,c('feature', 'model.Case.ControlCase Pr(>|t|)')]
  }else{
    aldex_sub <- aldex[,c('feature', 'model.Case.Control Pr(>|t|)')]
  }
  
  aldex_sub[,2] <- p.adjust(aldex_sub[,2], method="BH")
  
  ancom_sub <- ancom[,c('taxa_id', 'detected_0.8', 'detected_0.7')]
  rownames(ancom_sub) <- ancom_sub[,"taxa_id"]
  ancom_sub <- ancom_sub[intersect(feat_list,rownames(ancom_sub)),]
  
  ## have to clean up ancom_sub so that sig=0 and non-sig=1
  ancom_sub$detected_0.8 <- ifelse(ancom_sub$detected_0.8==TRUE, 0, 1)
  ancom_sub$detected_0.7 <- ifelse(ancom_sub$detected_0.7==TRUE, 0, 1)
  
  corncob_sub <- data.frame(feature=names(corncob$p),
                            p_fdr=corncob$p)
  
  rownames(corncob_sub) <- corncob_sub$feature
  
  corncob_sub <- corncob_sub[intersect(feat_list,rownames(corncob_sub)),]
  
  corncob_sub[,2] <- p.adjust(corncob_sub[,2], method="BH")
  ## fix corncob NAs and set them as 1 for our purpose... although this isn't totally correct...
  corncob_sub$p_fdr[is.na(corncob_sub$p_fdr)] <- 1

  maaslin2_p <- maaslin2[,c("feature", "pval")]
  rownames(maaslin2_p) <- maaslin2_p$feature
  maaslin2_p <- maaslin2_p[intersect(feat_list, rownames(maaslin2_p)),]
  maaslin2_p[,2] <- p.adjust(maaslin2_p[,2], method="BH")
  
  merge1 <- merge(x=aldex_sub, y=ancom_sub, by.x='feature', by.y='taxa_id', all=T)
  merge2 <- merge(x=merge1, y=corncob_sub, by.x='feature', by.y='feature', all=T)
  merge3 <- merge(x=merge2, y=maaslin2_p, by.x='feature', by.y='feature', all=T)
  merge3$detected_0.8[is.na(merge3$detected_0.8)] <- 1
  merge3$detected_0.7[is.na(merge3$detected_0.7)] <- 1
  merge3$pval[is.na(merge3$pval)] <- 1
  return(merge3)
  
}


#all the model estimate represent different "units" but we only care about sign in this case.
#in the case on ancom we will calculate the direction based on the difference in the mean CLR values
Comb_est_vals <- function(aldex, ancom, corncob, maaslin2, feat_list, clr_tab, meta, bin=T){
  
  aldex$feature <- rownames(aldex)
  aldex <- aldex[intersect(feat_list,rownames(aldex)),]
  ## we need to sort them all into the same order
  if(sum(grepl("ControlCase", colnames(aldex)))){
    aldex_sub <- aldex[,c('feature', 'model.Case.ControlCase Estimate')]
  }else{
    aldex_sub <- aldex[,c('feature', 'model.Case.Control Estimate')]
  }
  
  #ancom will need to be calculated from the CLR values as it does not by default output CLR values...
    
  clr_tab <- clr_tab[,feat_list]
  clr_tab$X <- rownames(clr_tab)
  
  clr_tab <- clr_tab %>% left_join(meta[,c("X", "Case.Control")])
  clr_tab <- clr_tab %>% group_by(Case.Control) %>% summarize_if(is.numeric, mean, na.rm=T)
  ancom_ests <- unlist(clr_tab[2,]) - unlist(clr_tab[1,])
  
  
  corncob_sub <- extract_CC_Est(corncob)
  rownames(corncob_sub) <- corncob_sub$feat
  corncob_sub <- corncob_sub[intersect(feat_list, rownames(corncob_sub)),]
  
  maaslin2_p <- maaslin2[,c("feature", "coef")]
  rownames(maaslin2_p) <- maaslin2_p$feature
  maaslin2_p <- maaslin2_p[intersect(feat_list, rownames(maaslin2_p)),]
  

  comb_tab <- aldex_sub %>% full_join(data.frame(feature=names(ancom_ests), ancom_est=ancom_ests))
  comb_tab <- comb_tab %>% full_join(corncob_sub, by="feature")
  comb_tab <- comb_tab %>% full_join(maaslin2_p, by="feature")
  
  if(bin){
    
    comb_tab$aldex_bin <- sapply(comb_tab[,2], bin_cov)
    comb_tab$ancom_bin <- sapply(comb_tab[,3], bin_cov)
    comb_tab$cc_bin <- sapply(comb_tab[,4], bin_cov)
    comb_tab$ma_bin <- sapply(comb_tab[,5], bin_cov)
  }
  
  comb_tab <- comb_tab %>% filter(feature != "Case.Control")
  return(comb_tab)
}


#convert values to + or -
#used for coloring the tool annotation
bin_cov <- function(x){
  if(is.na(x)){
    return(x)
  }
  
  if(x > 0){
    x <- 1
  }else if(x < 0){
    x <- -1
  }

  
  return(x)
}

#extract model estimates from the corncob object.
extract_CC_Est <- function(corncob){
  
  ret_df <- data.frame(feature=names(corncob$p))
  ret_df$cc_est <- NA
  for(i in 1:length(names(corncob$p))){
    if(is.na(corncob$all_models[[i]]))
      est_val <- NA
    else
      est_val <- corncob$all_models[[i]]$coefficients["mu.Case.ControlCase",1]
    
    ret_df[i,"cc_est"] <- est_val
  }
  
  return(ret_df)
}


## set non-sig estimates to 0
filt_non_sig_est <- function(pvals, ests, pval_thres){
  ret_ests <- ests
  stopifnot(identical(rownames(pvals), rownames(ests)))
  for(i in 1:nrow(pvals)){
    for(j in 2:ncol(pvals)){
      if(pvals[i,j] > pval_thres){
        ret_ests[i,j] <- 0
        ret_ests[i,j+4] <- 0
      }
    }
  }
  return(ret_ests)
}
```

## Plotting functions

```{r}
make_all_three_heat <- function(list_abund, list_pval, list_est, log=F, display_val=F){
  
  heat_list <- list()
  
  max_abund=-1
  
  for(i in names(list_abund)){
    temp_abund <- get_max_abund(list_abund[[i]], log)
    if(temp_abund > max_abund)
      max_abund <- temp_abund
  }
  
  
  if(log){
        for(i in names(list_abund)){
          heat_list[[i]] <- heatmap_function(abund=list_abund[[i]], pval = list_pval[[i]], est=list_est[[i]], 
                                             gradient = seq(-5, max_abund, by=0.1), main = i, show_row=F,
                                         show_legend = F, log=T, display_vals=display_val)
        }
  
  #draw labels
        labs_heat <- heatmap_function(abund=list_abund[[1]], pval = list_pval[[i]], est=list_est[[i]], 
                                      gradient = seq(-5, max_abund, by=0.1), main = "", show_row=T,
                                         show_legend = T, log=T, display_vals=display_val)
  }else{
      for(i in names(list_abund)){
        heat_list[[i]] <- heatmap_function(abund=list_abund[[i]], pval = list_pval[[i]], est=list_est[[i]], 
                                           gradient = seq(0, max_abund, by=0.001), main = i, show_row=F,
                                         show_legend = F, display_vals=display_val)
      }
  
  #draw labels
      labs_heat <- heatmap_function(abund=list_abund[[1]], pval = list_pval[[i]], est=list_est[[i]], 
                                    gradient = seq(0, max_abund, by=0.001), main = "", show_row=T,
                                         show_legend = T, display_vals=display_val)
  }

  # not the best way to do this...
  row_labs <- labs_heat$gtable
  #row_labs <- gtable_remove_grobs(row_labs, "main")
  row_labs <- gtable_remove_grobs(row_labs, "matrix")
  row_labs <- gtable_remove_grobs(row_labs, "col_names")
  row_labs <- gtable_remove_grobs(row_labs, "row_annotation")
  row_labs <- gtable_remove_grobs(row_labs, "row_annotation_names")
  
  final_plot <- plot_grid(heat_list[[1]]$gtable, heat_list[[2]]$gtable, heat_list[[3]]$gtable, row_labs,
                          ncol=4, rel_widths = c(1,1,1,2))
  return(final_plot)
}


heatmap_function <- function(abund, pval, est, gradient, main, show_row=T, show_legend=T, log=F, display_vals=F){
  abund <- abund %>% select(-age_diff, -shannon_entropy, -pielou_evenness, -observed_features, -faith_pd)
  mean_abund <- abund %>% group_by(Case.Control) %>% summarize(across(where(is.numeric), ~ mean(.x))) %>% as.data.frame()
  mean_abund[mean_abund==0] <- NA
  rownames(mean_abund) <- mean_abund[,1]
  mean_abund <- mean_abund[,-1]
  mean_abund_flip <- t(mean_abund)
  # if(log){
  #     message(max(log(mean_abund_flip)))
  # }else{
  #     message(max(mean_abund_flip))
  # }

  
  mean_abund_flip_full <- data.frame(mean_abund_flip)
  mean_abund_flip_full$feature <- rownames(mean_abund_flip)
  
  
  #full_data <- mean_abund_flip_full %>% left_join(pval)
  #full_data <- full_data %>% select(-detected_0.8) %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 1, .)))
  #full_data <- full_data %>% mutate_at(vars('model.Case.ControlCase Pr(>|t|)', 'detected_0.7', 'p_fdr', 'pval'),
                    #                                   function(x) ifelse(x <= 0.1, 1, 0))
  
  full_data <- mean_abund_flip_full %>% left_join(est)
  full_data <- full_data %>% mutate_if(is.numeric, funs(ifelse(is.na(.), 0, .)))
  
  
  rownames(full_data) <- full_data$feature
  
  
  colnames(full_data)[8:11] <- c("ALDEx2", "ANCOMII", "Corncob", "MaAsLin2")
  
  anno_cols <- list()
  ## possibly that all are postive or negative assocations...
  
  aldex2_range <- range(full_data$ALDEx2)
  ancom_range <- range(full_data$ANCOMII)
  corncob_range <- range(full_data$Corncob)
  maaslin_range <- range(full_data$MaAsLin2)
  
  if(identical(aldex2_range,c(-1, 1))){
    anno_cols[["ALDEx2"]] <- c("blue", "white", "red")
  }else if(identical(aldex2_range,c(-1, 0))){
        anno_cols[["ALDEx2"]] <- c("blue", "white")
  }else if(identical(aldex2_range,c(0, 0))){
        anno_cols[["ALDEx2"]] <- c("white")
  }else if(identical(aldex2_range,c(0, 1))){
        anno_cols[["ALDEx2"]] <- c("white", "red")
  }else if(identical(aldex2_range,c(-1,-1))){
        anno_cols[["ALDEx2"]] <- c("blue")
  }else if(identical(aldex2_range,c(1,1))){
        anno_cols[["ALDEx2"]] <- c("red")
  }

  if(identical(ancom_range,c(-1, 1))){
    anno_cols[["ANCOMII"]] <- c("blue", "white", "red")
  }else if(identical(ancom_range,c(-1, 0))){
        anno_cols[["ANCOMII"]] <- c("blue", "white")
  }else if(identical(ancom_range,c(0, 0))){
        anno_cols[["ANCOMII"]] <- c("white")
  }else if(identical(ancom_range,c(0, 1))){
        anno_cols[["ANCOMII"]] <- c("white", "red")
  }else if(identical(ancom_range,c(-1,-1))){
        anno_cols[["ANCOMII"]] <- c("blue")
  }else if(identical(ancom_range,c(1,1))){
        anno_cols[["ANCOMII"]] <- c("red")
  }
  
  if(identical(corncob_range,c(-1, 1))){
    anno_cols[["Corncob"]] <- c("blue", "white", "red")
  }else if(identical(corncob_range,c(-1, 0))){
        anno_cols[["Corncob"]] <- c("blue", "white")
  }else if(identical(corncob_range,c(0, 0))){
        anno_cols[["Corncob"]] <- c("white")
  }else if(identical(corncob_range,c(0, 1))){
        anno_cols[["Corncob"]] <- c("white")
  }else if(identical(corncob_range,c(-1,-1))){
        anno_cols[["Corncob"]] <- c("blue")
  }else if(identical(corncob_range,c(1,1))){
        anno_cols[["Corncob"]] <- c("red")
  }

  if(identical(maaslin_range,c(-1, 1))){
    anno_cols[["MaAsLin2"]] <- c("blue", "white", "red")
  }else if(identical(maaslin_range,c(-1, 0))){
        anno_cols[["MaAsLin2"]] <- c("blue", "white")
  }else if(identical(maaslin_range,c(0, 0))){
        anno_cols[["MaAsLin2"]] <- c("white")
  }else if(identical(maaslin_range,c(0, 1))){
        anno_cols[["MaAsLin2"]] <- c("white", "red")
  }else if(identical(maaslin_range,c(-1,-1))){
        anno_cols[["MaAsLin2"]] <- c("blue")
  }else if(identical(maaslin_range,c(1,1))){
        anno_cols[["MaAsLin2"]] <- c("red")
  }
  
  #return(list(full_data, anno_cols))
  if(!show_row){
    row_labs <- rep("", nrow(mean_abund_flip))
  }else{
   row_labs <- rownames(mean_abund_flip)
  }
  if(log){
      pheatmap(log10(mean_abund_flip),
         cluster_cols = F,
         cluster_rows = F,
         angle_col=90,
         annotation_row = full_data[,8:11],
         annotation_colors = anno_cols,
         annotation_legend = FALSE,
         main=main,
         color= colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(gradient)),
         breaks=gradient,
         labels_row = row_labs,
         legend=show_legend,
         display_numbers = display_vals,
         na_col = "grey")
    
  }else{
      pheatmap(mean_abund_flip,
         cluster_cols = F,
         cluster_rows = F,
         angle_col=90,
         annotation_row = full_data[,8:11],
         annotation_colors = anno_cols,
         annotation_legend = FALSE,
         main=main,
         color= colorRampPalette(rev(brewer.pal(n = 7, name = "RdYlBu")))(length(gradient)),
         breaks=gradient,
         labels_row = row_labs,
         legend=show_legend,
         display_number=display_vals,
         number_format = "%.1e",
         na_col="grey")
  }

}

get_max_abund <- function(x, log=F){
  x <- x %>% select(-age_diff, -shannon_entropy, -pielou_evenness, -observed_features, -faith_pd)
  mean_abund <- x %>% group_by(Case.Control) %>% summarize(across(where(is.numeric), ~ mean(.x))) %>% as.data.frame()
  
  rownames(mean_abund) <- mean_abund[,1]
  mean_abund <- mean_abund[,-1]
  mean_abund_flip <- t(mean_abund)
  if(log)
    return(max(log10(mean_abund_flip)))
  return(max(mean_abund_flip))
  
}

```


# Heat maps

## ASVs
```{r}
Breast_asvs <- rownames(ASV_Breast_filt)
Colon_asvs <- rownames(ASV_CRC_filt)
Prostate_asvs <- rownames(ASV_Prostate_filt)
```

### Breast
```{r}
Comb_Breast_ASV <- Comb_p_vals(aldex = Breast_ASV_aldex, 
                               ancom = Breast_ASV_ancom$out, 
                               corncob = Breast_ASV_corncob, 
                               maaslin2 = Breast_ASV_maaslin2,
                               feat_list = Breast_asvs)

rownames(Comb_Breast_ASV) <- Comb_Breast_ASV$feature

Comb_Breast_ASV_est <- Comb_est_vals(aldex=Breast_ASV_aldex,
                                     ancom=Breast_ASV_ancom,
                                     corncob=Breast_ASV_corncob,
                                     maaslin2=Breast_ASV_maaslin2,
                                     feat_list = Breast_asvs,
                                     ASV_Breast_CLR_flip,
                                     Breast_data,
                                     bin=T)

rownames(Comb_Breast_ASV_est) <- Comb_Breast_ASV_est$feature

Sig_Breast_ASV <- Comb_Breast_ASV %>% filter(`model.Case.ControlCase Pr(>|t|)` < 0.05 | pval < 0.05 | p_fdr < 0.05 | detected_0.7 < 0.05) %>%
  mutate_if(is.numeric, round, 4)

Sig_Breast_ASV_est <- Comb_Breast_ASV_est[rownames(Sig_Breast_ASV),]
Sig_Breast_ASV_est <- filt_non_sig_est(Sig_Breast_ASV[,-3], Sig_Breast_ASV_est, 0.05)

```

### Prostate
```{r}
Comb_Prostate_ASV <- Comb_p_vals(aldex=Prostate_ASV_aldex,
                                 ancom=Prostate_ASV_ancom$out,
                                 corncob = Prostate_ASV_corncob,
                                 maaslin2 = Prostate_ASV_maaslin2,
                                 feat_list = Prostate_asvs)

rownames(Comb_Prostate_ASV) <- Comb_Prostate_ASV$feature

Comb_Prostate_ASV_est <- Comb_est_vals(aldex=Prostate_ASV_aldex,
                                     ancom=Prostate_ASV_ancom,
                                     corncob=Prostate_ASV_corncob,
                                     maaslin2=Prostate_ASV_maaslin2,
                                     feat_list = Prostate_asvs,
                                     ASV_Prostate_CLR_flip,
                                     Prostate_data,
                                     bin=T)

rownames(Comb_Prostate_ASV_est) <- Comb_Prostate_ASV_est$feature

Sig_Prostate_ASV <- Comb_Prostate_ASV %>% filter(`model.Case.ControlCase Pr(>|t|)` < 0.05 | pval < .05 | p_fdr < .05 
                                                 | detected_0.7 < .05) %>% 
  mutate_if(is.numeric, round, 4)

Sig_Prostate_ASV_est <- Comb_Prostate_ASV_est[rownames(Sig_Prostate_ASV),]
Sig_Prostate_ASV_est <- filt_non_sig_est(Sig_Prostate_ASV[,-3], Sig_Prostate_ASV_est, 0.05)
```


### Colon
```{r}
Comb_Colon_ASV <- Comb_p_vals(aldex=Colon_ASV_aldex,
                              ancom=Colon_ASV_ancom$out,
                              corncob = Colon_ASV_corncob,
                              maaslin2 = Colon_ASV_maaslin2,
                              feat_list = Colon_asvs)

rownames(Comb_Colon_ASV) <- Comb_Colon_ASV$feature


Sig_Colon_ASV <- Comb_Colon_ASV %>% filter(`model.Case.ControlCase Pr(>|t|)` < .05 | pval < .05 | p_fdr < .05 
                                           | detected_0.7 < .05) %>%
  mutate_if(is.numeric, round, 4)



Comb_Colon_ASV_est <- Comb_est_vals(aldex=Colon_ASV_aldex,
                                     ancom=Colon_ASV_ancom,
                                     corncob=Colon_ASV_corncob,
                                     maaslin2=Colon_ASV_maaslin2,
                                     feat_list = Colon_asvs,
                                     ASV_CRC_CLR_flip,
                                     CRC_data,
                                     bin=T)

rownames(Comb_Colon_ASV_est) <- Comb_Colon_ASV_est$feature

Sig_Colon_ASV_est <- Comb_Colon_ASV_est[rownames(Sig_Colon_ASV),]
Sig_Colon_ASV_est <- filt_non_sig_est(Sig_Colon_ASV[,-3], Sig_Colon_ASV_est, 0.05)
```

### Combined
```{r}
all_sig_asvs <- c(Sig_Breast_ASV$feature, Sig_Prostate_ASV$feature, Sig_Colon_ASV$feature)
all_sig_asvs <- unique(all_sig_asvs)



#get prostate cancer abundances for sig taxa (RA)
Prostate_sig_abundances <- ASV_Prostate_RA_flip[,all_sig_asvs]
Prostate_sig_abundances$X <- rownames(Prostate_sig_abundances)

#get breast cancer abundances for sig taxa
Breast_sig_abundances <- ASV_Breast_RA_flip[,all_sig_asvs]
Breast_sig_abundances$X <- rownames(Breast_sig_abundances)

#get colon cancer abundances for sig taxa
Colon_sig_abundances <- ASV_CRC_RA_flip[,all_sig_asvs]
Colon_sig_abundances$X <- rownames(Colon_sig_abundances)
```

### Relabel taxa
```{r}

#read in ASV classifications so i can replace the labels
Taxa_classifications <- read.table("~/Private/Previous_Cancer_Project/RERUN_REV1/taxa/taxonomy/taxonomy.tsv", sep="\t",
                                   header=T, stringsAsFactors = F)

Taxa_classifications$Taxon <- gsub(".*__(?=.*__)", "", Taxa_classifications$Taxon, perl=T)
Taxa_classifications$Taxon <- make.unique(Taxa_classifications$Taxon)


## set up temporary dfs to store renamed labels
Rename_Prostate_sig_abundances <- Prostate_sig_abundances
Rename_Breast_sig_abundances <- Breast_sig_abundances
Rename_Colon_sig_abundances <- Colon_sig_abundances

#match asv col names to new taxa
colnames(Rename_Prostate_sig_abundances)[-ncol(Rename_Prostate_sig_abundances)] <- 
  Taxa_classifications$Taxon[match(colnames(Prostate_sig_abundances)[-ncol(Rename_Prostate_sig_abundances)], Taxa_classifications$Feature.ID)]

#see above
colnames(Rename_Breast_sig_abundances)[-ncol(Rename_Prostate_sig_abundances)] <- 
  Taxa_classifications$Taxon[match(colnames(Breast_sig_abundances)[-ncol(Rename_Prostate_sig_abundances)], Taxa_classifications$Feature.ID)]

#see above
colnames(Rename_Colon_sig_abundances)[-ncol(Rename_Colon_sig_abundances)] <- Taxa_classifications$Taxon[match(colnames(Colon_sig_abundances)[-ncol(Rename_Colon_sig_abundances)], Taxa_classifications$Feature.ID)]


#set up temporary dfs to hold sig taxa data
Rename_Sig_Breast_ASV <- Sig_Breast_ASV
Rename_Sig_Prostate_ASV <- Sig_Prostate_ASV
Rename_Sig_Colon_ASV <- Sig_Colon_ASV

#rename the p-value table
Rename_Sig_Breast_ASV$feature <- Taxa_classifications$Taxon[match(Sig_Breast_ASV$feature, Taxa_classifications$Feature.ID)]

Rename_Sig_Prostate_ASV$feature <- Taxa_classifications$Taxon[match(Sig_Prostate_ASV$feature, Taxa_classifications$Feature.ID)]

Rename_Sig_Colon_ASV$feature <- Taxa_classifications$Taxon[match(Sig_Colon_ASV$feature, Taxa_classifications$Feature.ID)]

#rename the est table
Rename_Sig_Breast_ASV_est <- Sig_Breast_ASV_est
Rename_Sig_Prostate_ASV_est <- Sig_Prostate_ASV_est
Rename_Sig_Colon_ASV_est <- Sig_Colon_ASV_est

Rename_Sig_Breast_ASV_est$feature <- Taxa_classifications$Taxon[match(Sig_Breast_ASV_est$feature, Taxa_classifications$Feature.ID)]

Rename_Sig_Prostate_ASV_est$feature <- Taxa_classifications$Taxon[match(Sig_Prostate_ASV_est$feature, Taxa_classifications$Feature.ID)]

Rename_Sig_Colon_ASV_est$feature <- Taxa_classifications$Taxon[match(Sig_Colon_ASV_est$feature, Taxa_classifications$Feature.ID)]
```

### Create heatmap
```{r}


Full_Prostate_data <- Prostate_data %>% full_join(Rename_Prostate_sig_abundances, by="X")

Full_Breast_Data <- Breast_data %>% full_join(Rename_Breast_sig_abundances, by="X")

Full_Colon_Data <- CRC_data %>% full_join(Rename_Colon_sig_abundances, by="X")



ASV_abund_list <- list(
  Breast=Full_Breast_Data,
  Prostate=Full_Prostate_data,
  Colon=Full_Colon_Data
)


ASV_p_list <- list(
  Breast=Rename_Sig_Breast_ASV,
  Prostate=Rename_Sig_Prostate_ASV,
  Colon=Rename_Sig_Colon_ASV
)

ASV_est_list <- list(
  Breast=Rename_Sig_Breast_ASV_est,
  Prostate=Rename_Sig_Prostate_ASV_est,
  Colon=Rename_Sig_Colon_ASV_est
)

test <- heatmap_function(abund=Full_Colon_Data, pval=Rename_Sig_Colon_ASV, est=Rename_Sig_Colon_ASV_est,
                         gradient = seq(-5, 0, by=0.1), main = "breast", show_row = T, show_legend = T, log = T, display_vals = T)

test

DA_heats <- make_all_three_heat(list_abund = ASV_abund_list, ASV_p_list, list_est=ASV_est_list, log=T, display_val = T)
DA_heats
```





## Genera

```{r}
Breast_genus <- rownames(Genus_Breast_filt)
Prostate_genus <- rownames(Genus_Prostate_filt)
Colon_genus <- rownames(Genus_CRC_filt)
```

### Breast

```{r}
Comb_Breast_Genus <- Comb_p_vals(aldex = Breast_GENUS_aldex,
                                 ancom = Breast_GENUS_ancom$out,
                                 corncob = Breast_GENUS_corncob,
                                 maaslin2 = Breast_GENUS_maaslin2,
                                 feat_list = Breast_genus)

rownames(Comb_Breast_Genus) <- Comb_Breast_Genus$feature

Sig_Breast_Genus <- Comb_Breast_Genus %>% filter(`model.Case.ControlCase Pr(>|t|)` < 0.05 | pval < 0.05 | p_fdr < 0.05 | detected_0.7 < 0.05) %>%
  mutate_if(is.numeric, round, 4)


Comb_Breast_Genus_est <- Comb_est_vals(aldex=Breast_GENUS_aldex,
                                       ancom=Breast_GENUS_ancom,
                                       corncob=Breast_GENUS_corncob,
                                       maaslin2 = Breast_GENUS_maaslin2,
                                       feat_list = Breast_genus,
                                       clr_tab = Genus_Breast_CLR_flip,
                                       meta=Breast_data,
                                       bin=T)

rownames(Comb_Breast_Genus_est) <- Comb_Breast_Genus_est$feature

Sig_Breast_Genus_est <- Comb_Breast_Genus_est[rownames(Sig_Breast_Genus),]
Sig_Breast_Genus_est <- filt_non_sig_est(Sig_Breast_Genus[,-3], Sig_Breast_Genus_est, 0.05)
```

### Prostate

```{r}
Comb_Prostate_GENUS <- Comb_p_vals(aldex=Prostate_GENUS_aldex,
                                   ancom = Prostate_GENUS_ancom$out,
                                   corncob = Prostate_GENUS_corncob,
                                   maaslin2 = Prostate_GENUS_maaslin2,
                                   feat_list = Prostate_genus)

rownames(Comb_Prostate_GENUS) <- Comb_Prostate_GENUS$feature

Sig_Prostate_Genus <- Comb_Prostate_GENUS %>% filter(`model.Case.ControlCase Pr(>|t|)` < .05 | pval < .05 | p_fdr < .05| detected_0.7 < .05) %>%
  mutate_if(is.numeric, round, 4)

Comb_Prostate_Genus_est <- Comb_est_vals(aldex=Prostate_GENUS_aldex,
                                         ancom=Prostate_GENUS_ancom,
                                         corncob = Prostate_GENUS_corncob,
                                         maaslin2 = Prostate_GENUS_maaslin2,
                                         feat_list = Prostate_genus,
                                         clr_tab = Genus_Prostate_CLR_flip,
                                         meta=Prostate_data,
                                         bin=T)

rownames(Comb_Prostate_Genus_est) <- Comb_Prostate_Genus_est$feature


Sig_Prostate_Genus_est <- Comb_Prostate_Genus_est[rownames(Sig_Prostate_Genus),]
Sig_Prostate_Genus_est <- filt_non_sig_est(Sig_Prostate_Genus[,-3], Sig_Prostate_Genus_est, 0.05)
```

### Colon

```{r}
Comb_Colon_GENUS <- Comb_p_vals(aldex=Colon_GENUS_aldex,
                                ancom = Colon_GENUS_ancom$out,
                                corncob = Colon_GENUS_corncob,
                                maaslin2 = Colon_GENUS_maaslin2,
                                feat_list = Colon_genus)

rownames(Comb_Colon_GENUS) <- Comb_Colon_GENUS$feature

Sig_Colon_Genus <- Comb_Colon_GENUS %>% filter(`model.Case.ControlCase Pr(>|t|)` < .05 | pval < .05 | p_fdr < .05| detected_0.7 < .05) %>%
  mutate_if(is.numeric, round, 4)

Comb_Colon_GENUS_est <- Comb_est_vals(aldex=Colon_GENUS_aldex,
                                      ancom=Colon_GENUS_ancom,
                                      corncob = Colon_GENUS_corncob,
                                      maaslin2 = Colon_GENUS_maaslin2,
                                      feat_list = Colon_genus,
                                      clr_tab=Genus_CRC_CLR_flip,
                                      meta=CRC_data,
                                      bin=T)


rownames(Comb_Colon_GENUS_est) <- Comb_Colon_GENUS_est$feature

Sig_Colon_Genus_est <- Comb_Colon_GENUS_est[rownames(Sig_Colon_Genus),]
Sig_Colon_Genus_est <- filt_non_sig_est(Sig_Colon_Genus[,-3], Sig_Colon_Genus_est, 0.05)
```

### combined

```{r}
all_sig_genera <- c(Sig_Breast_Genus$feature, Sig_Prostate_Genus$feature, Sig_Colon_Genus$feature)
all_sig_genera <- unique(all_sig_genera)


Prostate_sig_abundances_genera <- Genus_Prostate_RA_flip[,all_sig_genera]
Prostate_sig_abundances_genera$X <- rownames(Prostate_sig_abundances_genera)


Breast_sig_abundances_genera <- Genus_Breast_RA_flip[,all_sig_genera]
Breast_sig_abundances_genera$X <- rownames(Breast_sig_abundances)


Colon_sig_abundances_genera <- Genus_CRC_RA_flip[,all_sig_genera]
Colon_sig_abundances_genera$X <- rownames(Colon_sig_abundances_genera)
```

### fix label names
```{r}
Full_Prostate_data_genera <- Prostate_data %>% full_join(Prostate_sig_abundances_genera, by="X")
Full_Breast_data_genera <- Breast_data %>% full_join(Breast_sig_abundances_genera, by="X")
Full_Colon_data_genera <- CRC_data %>% full_join(Colon_sig_abundances_genera, by="X")


#fix Genera name
colnames(Full_Breast_data_genera) <- gsub(".*__(?=.*__)", "", colnames(Full_Breast_data_genera), perl=T)
colnames(Full_Prostate_data_genera) <- gsub(".*__(?=.*__)", "", colnames(Full_Prostate_data_genera), perl=T)
colnames(Full_Colon_data_genera) <- gsub(".*__(?=.*__)", "", colnames(Full_Colon_data_genera), perl=T)

Sig_Breast_Genus$feature <- gsub(".*__(?=.*__)", "", Sig_Breast_Genus$feature, perl=T)
Sig_Prostate_Genus$feature <- gsub(".*__(?=.*__)", "", Sig_Prostate_Genus$feature, perl=T)
Sig_Colon_Genus$feature <- gsub(".*__(?=.*__)", "", Sig_Colon_Genus$feature, perl=T)

Sig_Breast_Genus_est$feature <- gsub(".*__(?=.*__)", "", Sig_Breast_Genus_est$feature, perl=T)
Sig_Prostate_Genus_est$feature <- gsub(".*__(?=.*__)", "", Sig_Prostate_Genus_est$feature, perl=T)
Sig_Colon_Genus_est$feature <- gsub(".*__(?=.*__)", "", Sig_Colon_Genus_est$feature, perl=T)
```

### Plot

```{r}
Genus_abund_list <- list(
  Breast=Full_Breast_data_genera,
  Prostate=Full_Prostate_data_genera,
  Colon=Full_Colon_data_genera
)

Genus_p_list <- list(
  Breast=Sig_Breast_Genus,
  Prostate=Sig_Prostate_Genus,
  Colon=Sig_Colon_Genus
)

Genus_est_list <- list(
  Breast=Sig_Breast_Genus_est,
  Prostate=Sig_Prostate_Genus_est,
  Colon=Sig_Colon_Genus_est
)

Genus_DA_heat <- make_all_three_heat(list_abund = Genus_abund_list, list_pval = Genus_p_list, list_est = Genus_est_list, log = T, display_val = T)
Genus_DA_heat
```