---
title: "Reterospective Pub Figs"
author: "Jacob T. Nearing"
date: "3/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(broom)
library(kableExtra)
library(ggplot2)
library(ggbeeswarm)
library(cowplot)

library(reshape2)
library(ggpubr)
library(rstatix)
library(vegan)

theme_set(theme_cowplot())

#already tests
remove_non_matching_caseIDS <- function(df){
  
    ## remove controls that don;t have associated case samples b/c they were removed due to poor sequencing
    df_cases <- df %>% filter(Case.Control=="Case")
    CaseIDs <- df_cases$CaseID
    df_filt <- df %>% filter(CaseID %in% CaseIDs)

    ## remove cases that don't have any controls associtated because they were removed due to poor sequencing 
    ## these cases can only happen when the total number of samples with that CASEID is 1 
    Summ_data <- df_filt %>% group_by(CaseID) %>% summarize(n())
    remove_cases <- Summ_data$CaseID[which(Summ_data$`n()`==1)]
    
    if(length(remove_cases!=0)){
      message(paste0("removing ", remove_cases, " Case samples due to no controls"))
      test <- df_filt[-which(df_filt$CaseID %in% remove_cases),]
      return(test)  
    }

    
    return(df_filt)
  
}

#already tested
match_age_diff <- function(df){
  
  case_df <- df %>% filter(Case.Control=="Case")
  for(i in 1:nrow(df)){
    
    if(is.na(df[i,'age_diff']) & df[i,'Case.Control']=="Control"){
      temp_caseID <- df[i,'CaseID']
      message(temp_caseID)
      temp_age_diff <- case_df$age_diff[which(case_df$CaseID==temp_caseID)]
      message(temp_age_diff)
      df[i,'age_diff'] <- temp_age_diff
    }
  }
  return(df)
}

run_lm_comb_data <- function(data, type_split, main_variable, covariates, estimate_var){
  
  anova_res <- list()
  ggpubr_res <- list()
  
  for(i in unique(data[,type_split])){
    pass_in_data <- data %>% filter(!!sym(type_split)==i)
    message(type_split)
    res <- run_lm_add_using_ggplot(pass_in_data, main_variable = main_variable, covariates = covariates,
                            estimate_var = estimate_var, fct=i, type_split)
    anova_res[[i]] <- res[[1]]
    ggpubr_res[[i]] <- res[[2]]
  }
  ggpurb_df <- do.call(rbind, ggpubr_res)
  
  return(list(anova_res, ggpurb_df))
  
}

data("USMortality")

test_lm_comb <- run_lm_comb_data(USMortality, type_split="Status", main_variable = "Sex", covariates = NULL, estimate_var = "Rate")
test_lm_comb

## manually test

Rural <- USMortality %>% filter(Status=="Rural")
anova(lm(Rate ~ Sex, data=Rural))
#checks out

Urban <- USMortality %>% filter(Status=="Urban")
anova(lm(Rate ~ Sex, data=Urban))
#Checks out

run_lm_add_using_ggplot <- function(data, main_variable, covariates, estimate_var, fct, type_split){
    
  
    formula <- paste(covariates, main_variable, sep=" + ")
    final_formula <- paste(estimate_var, "~", formula, sep=" ")
    message(final_formula)
    model <- lm(final_formula, data)
    anova_res <- anova(model)
    anova_res$feat <- rownames(anova_res)
    anova_filt <- anova_res %>% filter(feat==main_variable)
    pval <- anova_filt$`Pr(>F)`
    
    if(is.null(fct)){
      ggpurb_format <- data.frame(group1=levels(data[,main_variable])[1],
                                group2=levels(data[,main_variable])[2],
                                p.adj=pval)
    }else{
      ggpurb_format <- data.frame(group1=levels(data[,main_variable])[1],
                                group2=levels(data[,main_variable])[2],
                                p.adj=round(pval,3),
                                supp=fct,
                                .y.=max(data[,estimate_var])*1.20)
      
      colnames(ggpurb_format)[4] <- type_split
      
    }
    
    return(list(anova_res, ggpurb_format))
}


make_alpha_div_plot <- function(df, split_type, main_variable, covariates, estimate_var,ylab,pval=T,hide.ns){
  
  stats <- run_lm_comb_data(data=df, type_split = split_type, main_variable = main_variable,
                                        covariates = covariates, estimate_var = estimate_var)
  
  plot <- ggplot(df, aes_string(x=main_variable, y=estimate_var)) +
  geom_boxplot(alpha=0.1, width=0.1, outlier.alpha = 0) +
  geom_violin(alpha=0.2, aes_string(fill=main_variable)) +
  geom_quasirandom(alpha=0.2) +
  xlab("") +
  ylab(ylab) +
  facet_grid(reformulate(split_type))

  if(pval==T){
    plot <- plot + stat_pvalue_manual(stats[[2]], label="p.adj", hide.ns=hide.ns, y.position = '.y.')
  }
  
  return(plot)
}




make_all_alpha_div_plots <- function(df, split_type, main_variable, covariates, estimate_var, ylab, pval=T, hide.ns=T){
  plot_list <- list()
  j <- 1
  for(i in ylab){
    plot_list[[i]] <- make_alpha_div_plot(df, split_type, main_variable, covariates, estimate_var[j], i, pval,hide.ns)
    j <- j + 1
  }
  return(plot_list)
  
}

test_vals <- c("Rate", "SE")

test_plots <- make_all_alpha_div_plots(USMortality, split_type = "Status", main_variable = "Sex", covariates = NULL, estimate_var = test_vals,
                                       ylab=c("Rate", "SE"))

test_plots[[1]]
test_plots[[2]]


#all functions check out based on my unit tests

```

# Alpha Diversity Figures

```{r load data}

Metadata_Breast <- read.csv("~/Private/Previous_Cancer_Project/Metadata/Metadata_Breast_21_03_11.csv", row.names = 1)
Metadata_CRC <- read.csv("~/Private/Previous_Cancer_Project/Metadata/Metadata_CRC_21_03_11.csv", row.names = 1)
Metadata_Prostate <- read.csv("~/Private/Previous_Cancer_Project/Metadata/Metadata_Prostate_21_03_11.csv", row.names = 1)


### Load in alpha data

Alpha_Breast <- readRDS("~/Private/Previous_Cancer_Project/Data/Alpha_data_Breast.RDS")
Alpha_CRC <- readRDS("~/Private/Previous_Cancer_Project/Data/Alpha_data_CRC.RDS")
Alpha_Prostate <- readRDS("~/Private/Previous_Cancer_Project/Data/Alpha_data_prostate.RDS")


Metadata_Breast <- Metadata_Breast[rownames(Alpha_Breast),]
Metadata_CRC <- Metadata_CRC[rownames(Alpha_CRC),]
Metadata_Prostate <- Metadata_Prostate[rownames(Alpha_Prostate),]

Metadata_Breast_Cases <- Metadata_Breast[which(Metadata_Breast$Case.Control==1),]
Metadata_CRC_Cases <- Metadata_CRC[which(Metadata_CRC$Case.Control==1),]
Metadata_Prostate_Cases <- Metadata_Prostate[which(Metadata_Prostate$Case.Control==1),]

Alpha_Breast_Cases <- Alpha_Breast[rownames(Metadata_Breast_Cases),]
Alpha_CRC_Cases <- Alpha_CRC[rownames(Metadata_CRC_Cases),]
Alpha_Prostate_Cases <- Alpha_Prostate[rownames(Metadata_Prostate_Cases),]

Alpha_Breast$X <- rownames(Alpha_Breast)
Breast_data <- Alpha_Breast %>% left_join(Metadata_Breast, by="X")
Breast_data$Case.Control <- ifelse(Breast_data$Case.Control==1, "Case", "Control")

Alpha_Prostate$X <- rownames(Alpha_Prostate)
Prostate_data <- Alpha_Prostate %>% left_join(Metadata_Prostate, by="X")
Prostate_data$Case.Control <- ifelse(Prostate_data$Case.Control==1, "Case", "Control")

Alpha_CRC$X <- rownames(Alpha_CRC)
CRC_data <- Alpha_CRC %>% left_join(Metadata_CRC, by="X")
CRC_data$Case.Control <- ifelse(CRC_data$Case.Control==1, "Case", "Control")

Alpha_Breast_Cases$X <- rownames(Alpha_Breast_Cases)
Breast_case_data <- Alpha_Breast_Cases %>% left_join(Metadata_Breast_Cases, by="X")
Breast_case_data$Case.Control <- ifelse(Breast_case_data$Case.Control==1, "Case", "Control")

Alpha_Prostate_Cases$X <- rownames(Alpha_Prostate_Cases)
Prostate_case_data <- Alpha_Prostate_Cases %>% left_join(Metadata_Prostate_Cases, by="X")
Prostate_case_data$Case.Control <- ifelse(Prostate_case_data$Case.Control==1, "Case", "Control")

Alpha_CRC_Cases$X <- rownames(Alpha_CRC_Cases)
CRC_case_data <- Alpha_CRC_Cases %>% left_join(Metadata_CRC_Cases, by="X")
CRC_case_data$Case.Control <- ifelse(CRC_case_data$Case.Control==1, "Case", "Control")


Breast_data$age_diff <- Breast_data$YEARS_SINCE_BC
Prostate_data$age_diff <- Prostate_data$YEARS_SINCE_PC
CRC_data$age_diff <- CRC_data$YEARS_SINCE_CRC_CANCER

Breast_data$Type <- "Breast"
Prostate_data$Type <- "Prostate"
CRC_data$Type <- "Colon"

Breast_data$Case.Control <- factor(Breast_data$Case.Control, levels=c("Control", "Case"))
Prostate_data$Case.Control <- factor(Prostate_data$Case.Control, levels=c("Control", "Case"))
CRC_data$Case.Control <- factor(CRC_data$Case.Control, levels=c("Control","Case"))

Breast_data <- remove_non_matching_caseIDS(Breast_data)
Prostate_data <- remove_non_matching_caseIDS(Prostate_data)
CRC_data <- remove_non_matching_caseIDS(CRC_data)

Breast_data_comb <- Breast_data %>% select(faith_pd, shannon, pielou_e, observed_otus, X, Case.Control, age_diff, Type, CaseID, Extraction_Number)
Prostate_data_comb <- Prostate_data %>% select(faith_pd, shannon, pielou_e, observed_otus, X, Case.Control, age_diff, Type, CaseID, Extraction_Number)
CRC_data_comb <- CRC_data %>% select(faith_pd, shannon, pielou_e, observed_otus, X, Case.Control, age_diff, Type, CaseID, Extraction_Number)


Combined_data <- rbind(Breast_data_comb, Prostate_data_comb, CRC_data_comb)


Melted_Combined_data <- melt(Combined_data, id.vars=c("X", "Case.Control", "age_diff", "Type"), 
                             measure.vars = c("faith_pd", "shannon", "pielou_e", "observed_otus"),
                             variable.name="Alpha_Metric")


Case_data_filt <- Combined_data %>% filter(age_diff <= 6)

Case_time_keep <- Case_data_filt$CaseID

Combined_data_filt <- Combined_data %>% filter(CaseID %in% Case_time_keep)

```


## Supplemental Figure


### No Time Filter

```{r}

est_vars <- c("faith_pd", "observed_otus", "shannon", "pielou_e")
ylabs <- c("Faith's PD", "Richness", "Shannon Div.", "Evenness")


all_cancer_plots <- make_all_alpha_div_plots(df=Combined_data, split_type = "Type", main_variable = "Case.Control",
                         covariates = "Extraction_Number", estimate_var = est_vars, ylab = ylabs, pval=T, hide.ns=F)

#verify p-values
test <- Combined_data %>% filter(Type=="Breast")
summary(lm(faith_pd ~ Case.Control + Extraction_Number, data=test))
#checks out.
summary(lm(observed_otus ~ Case.Control + Extraction_Number, data=test))
#checks out
test <- Combined_data %>% filter(Type=="Colon")
summary(lm(faith_pd ~ Case.Control + Extraction_Number, data=test))
#checks out.
summary(lm(observed_otus ~ Case.Control + Extraction_Number, data=test))
#all checks outs.


saveRDS(all_cancer_plots, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/ret_path_alpha_all.RDS")

all_cancer_plots_no_pval <- make_all_alpha_div_plots(df=Combined_data, split_type = "Type", main_variable = "Case.Control",
                         covariates = "Extraction_Number", estimate_var = est_vars, ylab = ylabs, pval=F)

saveRDS(all_cancer_plots_no_pval, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/ret_path_alpha_all_no_p.RDS") 

```


### Time Filtering

```{r}
#checks out.
range(Combined_data_filt$age_diff, na.rm=T)

all_cancer_plots_filt_6 <- make_all_alpha_div_plots(df=Combined_data_filt, split_type = "Type", main_variable = "Case.Control",
                         covariates = "Extraction_Number", estimate_var = est_vars, ylab = ylabs, pval=T, hide.ns=T)

saveRDS(all_cancer_plots_filt_6, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/ret_path_alpha_all_time6.RDS")

all_cancer_plots_filt_6_no_p <- make_all_alpha_div_plots(df=Combined_data_filt, split_type = "Type", main_variable = "Case.Control",
                         covariates = "Extraction_Number", estimate_var = est_vars, ylab = ylabs, pval=F)

saveRDS(all_cancer_plots_filt_6_no_p, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/ret_path_all_time6_no_p.RDS")
```


# Beta Diversity Figures

```{r}
plot_pca_all <- function(dist, meta, title, facet_type=NULL){
  
  dist_filt <- dist[meta$X, meta$X]
  
  dist_pca <- cmdscale(dist_filt, eig=T)
  
  dist_plot_data <- data.frame(PC1=dist_pca$points[,1],
                               PC2=dist_pca$points[,2],
                               Case.Control=meta$Case.Control_type,
                               facet=meta[,facet_type])
  
  
  PC1_size <- round(dist_pca$eig[1]/sum(dist_pca$eig)*100,1)
  PC2_size <- round(dist_pca$eig[2]/sum(dist_pca$eig)*100,1)
  
  PC1_lab <- paste0("PC1 ", PC1_size, "%")
  PC2_lab <- paste0("PC2 ", PC2_size, "%")
  
  if(!is.null(facet_type)){
    dist_plot <- ggplot(dist_plot_data, aes(x=PC1, y=PC2)) +
      geom_point(aes(fill=Case.Control, color=Case.Control)) +
      stat_ellipse(aes(group=Case.Control, color=Case.Control)) +
      theme(legend.title = element_blank()) +
      xlab(PC1_lab) +
      ylab(PC2_lab) +
      ggtitle(title) +
      facet_grid(~ facet)
    
    return(dist_plot)
    
  }else{
      
  dist_plot <- ggplot(dist_plot_data, aes(x=PC1, y=PC2, fill=Case.Control, group=Case.Control, color=Case.Control)) +
    geom_point() +
    stat_ellipse() +
    theme(legend.title = element_blank()) +
    xlab(PC1_lab) +
    ylab(PC2_lab) +
    ggtitle(title)
  
  return(dist_plot)
    
  }

}


## testing
##make fake distance matrix

test_dist <- data.frame(T1 = c(0,.8,.6,.4,.2),
                        T2 = c(.8,0,.8,.6,.4),
                        T3 = c(.6,.8,0,.8,.6),
                        T4 = c(.4,.6,.8,0,.8),
                        T5 = c(.2,.4,.6,.8,0))

test_names <- c("T1", "T2", "T3", "T4", "T5")
colnames(test_dist) <- test_names
rownames(test_dist) <- test_names
test_metadata <- data.frame(Case.Control_type=c("Cancer","Cancer","Healthy","Healthy","Cancer"),
                            X=test_names)
test_metadata$Case.Control <- test_metadata$Case.Control_type
rownames(test_metadata) <- test_names

plot_pca_all(dist=test_dist, meta = test_metadata, title = "test case", facet_type = NULL)
#looks like it works.

                       

run_adonis_by_facet <- function(dist, meta, facet, covariates, main_variable){
  
  formula <- paste(covariates, main_variable, sep=" + ")
  final_formula <- as.formula(paste("temp_dist", formula, sep="~"))
  message(final_formula)
  message(facet)
  meta_splits <- group_split(meta, !!ensym(facet))
  adonis_res <- list()

  for(i in 1:length(meta_splits)){
    name <- as.character(meta_splits[[i]][1,facet,drop=T])
    message(name)
    samps <- as.data.frame(meta_splits[[i]][,"X"])
    temp_dist <- dist[samps$X, samps$X]
    message("running adonis")
    
    temp_res<- adonis(formula=final_formula, data=meta_splits[[i]], by="margin")
    temp_res <- as.data.frame(temp_res$aov.tab)
    temp_res$feature <- rownames(temp_res)
    
    adonis_res[[name]] <- temp_res
  }
  
  main_vars_p <- c()
  for(i in 1:length(adonis_res)){
    temp_p <- adonis_res[[i]] %>% filter(feature==main_variable) %>% select(`Pr(>F)`)
    main_vars_p <- c(main_vars_p, paste("p: ",as.character(temp_p), sep=""))
  }
  
  annotation_df <- data.frame(col1=levels(meta[,facet]),
                              label=main_vars_p)
  colnames(annotation_df)[1] <- "facet"
  return(list(annotation_df,adonis_res))
}

## manually test this below


make_all_plots_and_adonis <- function(list_dist, list_title, metadata, facet, covariate, main_variable){
  
  pca_plots <- list()
  adonis_res <- list()
  
  for(i in names(list_dist)){
    pca_plots[[i]] <- plot_pca_all(dist=list_dist[[i]], meta = metadata, title = list_title[[i]], facet_type = facet)
    adonis_res[[i]] <- run_adonis_by_facet(dist=list_dist[[i]], meta=metadata, facet = facet, covariates = covariate, main_variable = main_variable) 
  }
  
  return(list(pca_plots, adonis_res))
}


## test


```

## All cancers in one

```{r}
WEIGHTED_UNIFRAC <- "~/Private/Sequences/Redo_Combined_Data/deblur/core_div_metrics_5000/beta_data_5000/weighted_unifrac_distance_matrix.tsv"
BRAY_CURTIS <- "~/Private/Sequences/Redo_Combined_Data/deblur/core_div_metrics_5000/beta_data_5000/bray_curtis_distance_matrix.tsv"
UNWEIGHTED_UNIFRAC <- "~/Private/Sequences/Redo_Combined_Data/deblur/core_div_metrics_5000/beta_data_5000/unweighted_unifrac_distance_matrix.tsv"

w_unifrac <- read.table(WEIGHTED_UNIFRAC, sep="\t", header=T, row.names = 1, comment.char="", quote="", check.names = F)

bray_curt <- read.table(BRAY_CURTIS, sep="\t", header=T, row.names = 1, comment.char = "", quote="", check.names = F)

u_unifrac <- read.table(UNWEIGHTED_UNIFRAC, sep="\t", header=T, row.names = 1, comment.char="", quote="", check.names = F)

Combined_data$Case.Control_type <- as.character(Combined_data$Case.Control)

Combined_data$Case.Control_type[which(Combined_data$Case.Control_type=="Case")] <- Combined_data$Type[which(Combined_data$Case.Control=="Case")]


Combined_data_filt$Case.Control_type <- as.character(Combined_data_filt$Case.Control)

Combined_data_filt$Case.Control_type[which(Combined_data_filt$Case.Control_type=="Case")] <- Combined_data_filt$Type[which(Combined_data_filt$Case.Control=="Case")]
```

### non-time-filt

```{r}
weighted_uni_comb <- w_unifrac[Combined_data$X, Combined_data$X]
unweighted_uni_comb <- u_unifrac[Combined_data$X, Combined_data$X]
bray_comb <- bray_curt[Combined_data$X, Combined_data$X]
```




```{r}
weighted_comb_plot <- plot_pca_all(weighted_uni_comb, Combined_data, "Weighted UniFrac")

saveRDS(weighted_comb_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_wuni.RDS")

unweighted_comb_plot <- plot_pca_all(unweighted_uni_comb, Combined_data, "Unweighted UniFrac")

saveRDS(unweighted_comb_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_uuni.RDS")

bray_comb_plot <- plot_pca_all(bray_comb, Combined_data, "Bray Curtis")

saveRDS(bray_comb_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_bray.RDS")

```

# adonis test comparing all samples between each other
```{r}
library(vegan)
adonis(weighted_uni_comb ~ Combined_data$Extraction_Number + Combined_data$Case.Control_type)
adonis(bray_comb ~ Combined_data$Extraction_Number + Combined_data$Case.Control_type)
adonis(unweighted_uni_comb ~ Combined_data$Extraction_Number + Combined_data$Case.Control_type)

#compare Colon vs. controls regardless of matching.
pair_crc <- Combined_data %>% filter(Combined_data$Case.Control_type %in% c("Colon", "Control"))
pair_unweight <- unweighted_uni_comb[pair_crc$X, pair_crc$X]

pair_breast <- Combined_data %>% filter(Combined_data$Case.Control_type %in% c("Breast", "Control"))
pair_unweight_br <- unweighted_uni_comb[pair_breast$X, pair_breast$X] 
  
  
adonis(pair_unweight ~ pair_crc$Extraction_Number + pair_crc$Case.Control_type)
adonis(pair_unweight_br ~ pair_breast$Extraction_Number + pair_breast$Case.Control_type)
```

### time_filt

This figures were not included in the final manuscript as they do not show any major differences from the non-time filtered figures and were fairly repeatitive. 

```{r}
weighted_comb_time_plot <- plot_pca_all(weighted_uni_comb, Combined_data_filt, "Weighted UniFrac")

saveRDS(weighted_comb_time_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_wuni_time.RDS")

unweighted_comb_time_plot <- plot_pca_all(unweighted_uni_comb, Combined_data_filt, "Unweighted UniFrac")

saveRDS(unweighted_comb_time_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_uuni_time.RDS")

bray_comb_time_plot <- plot_pca_all(bray_comb, Combined_data_filt, "Bray Curtis")

saveRDS(bray_comb_time_plot, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Ret_PATH_bray_time.RDS")
```


## Individual

### no time filt

#### Breast
```{r}
Breast_data_comb <- Combined_data %>% filter(Type=="Breast")
Breast_data_comb$Case.Control_type <- Breast_data_comb$Case.Control
Breast_data_comb$Type <- factor(Breast_data_comb$Type)

dist_list <- list(
  'W_uni' = weighted_uni_comb,
  'U_uni' =  unweighted_uni_comb,
  'bray' = bray_comb
)

title_list <- list(
  'W_uni' = "Weighted UniFrac",
  'U_uni' = "Unweighted UniFrac",
  'bray' = "Bray Curtis"
)


Breast_plots <- make_all_plots_and_adonis(list_dist = dist_list, list_title = title_list, metadata = Breast_data_comb, facet = "Type", covariate = "Extraction_Number", main_variable = "Case.Control")

#double check some of the adonis test results to make sure they are right.

Br_w_uni <- weighted_uni_comb[Breast_data_comb$X, Breast_data_comb$X]
ad_test <- adonis2(Br_w_uni ~ Breast_data_comb$Extraction_Number + Breast_data_comb$Case.Control)
ad_test
#match up

saveRDS(Breast_plots, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Breast_plots_all_beta.RDS")

```

#### Prostate
```{r}
Prostate_data_comb <- Combined_data %>% filter(Type=="Prostate")
Prostate_data_comb$Case.Control_type <- Prostate_data_comb$Case.Control
Prostate_data_comb$Type <- factor(Prostate_data_comb$Type)

Prostate_plots <- make_all_plots_and_adonis(list_dist=dist_list, list_title = title_list, metadata=Prostate_data_comb, 
                                            facet="Type", covariate = "Extraction_Number",
                                            main_variable = "Case.Control")


#double check some values
Pr_u_uni <- unweighted_uni_comb[Prostate_data_comb$X, Prostate_data_comb$X]
Pr_test <- adonis2(Pr_u_uni ~ Prostate_data_comb$Extraction_Number + Prostate_data_comb$Case.Control)
Pr_test

#matches up

saveRDS(Prostate_plots, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Prostate_plots_all_beta.RDS")
```


#### CRC
```{r}
CRC_data_comb <- Combined_data %>% filter(Type=="Colon")
CRC_data_comb$Case.Control_type <- CRC_data_comb$Case.Control
CRC_data_comb$Type <- factor(CRC_data_comb$Type)

CRC_plots <- make_all_plots_and_adonis(list_dist = dist_list, list_title = title_list, metadata=CRC_data_comb, facet="Type", covariate = "Extraction_Number",
                                       main_variable = "Case.Control")
saveRDS(CRC_plots, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/CRC_plots_all_beta.RDS")
```

### Time filt

shows similiar results so was not included in final manuscript.

#### Breast
```{r}
Breast_data_comb <- Combined_data_filt %>% filter(Type=="Breast")
Breast_data_comb$Case.Control_type <- Breast_data_comb$Case.Control
Breast_data_comb$Type <- factor(Breast_data_comb$Type)

Breast_plots_time <- make_all_plots_and_adonis(list_dist = dist_list, list_title = title_list, metadata=Breast_data_comb, facet="Type", covariate = "Extraction_Number",
                                               main_variable = "Case.Control")

saveRDS(Breast_plots_time, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Breast_plots_all_beta_time6.RDS")
```


#### Prostate
```{r}
Prostate_data_comb <- Combined_data_filt %>% filter(Type=="Prostate")
Prostate_data_comb$Case.Control_type <- Prostate_data_comb$Case.Control_type
Prostate_data_comb$Type <- factor(Prostate_data_comb$Type)

Prostate_plots_times <- make_all_plots_and_adonis(list_dist = dist_list, list_title = title_list, metadata = Prostate_data_comb, facet="Type", covariate = "Extraction_Number",
                                                  main_variable="Case.Control")

saveRDS(Prostate_plots_times, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/Prostate_plots_all_beta_time6.RDS")
```


#### CRC
```{r}
CRC_data_comb <- Combined_data_filt %>% filter(Type=="Colon")
CRC_data_comb$Case.Control_type <- CRC_data_comb$Case.Control
CRC_data_comb$Type <- factor(CRC_data_comb$Type)

CRC_plots_times <- make_all_plots_and_adonis(list_dist = dist_list, list_title = title_list, metadata=CRC_data_comb, facet="Type", covariate = "Extraction_Number",
                                             main_variable = "Case.Control")

saveRDS(CRC_plots_times, "~/Private/CHAPTER_4/Publication_Markdowns/Display_items/intermediate_RDS2/CRC_plots_all_beta_time6.RDS")
```

# alpha div correlation with time


```{r}
Combined_data %>% filter(Case.Control=="Control") %>% summarize(mean(faith_pd))

faiths_cor_data <- ggplot(Combined_data, aes(x=age_diff, y=faith_pd)) + geom_point() +
  facet_grid(~ Type) + geom_line(y=13.26, color="red", linetype="dashed") +
  ylab("Faith's Phylogenetic Diversity") + xlab("Time after diagnosis")
faiths_cor_data

Combined_data %>% filter(Case.Control=="Control") %>% summarize(mean(observed_otus))

richness_cor_data <- ggplot(Combined_data, aes(x=age_diff, y=observed_otus)) + geom_point() +
  facet_grid(~ Type) + geom_line(y=128.87, color="red", linetype="dashed") +
  ylab("Richness") + xlab("Time after diagnosis")

Combined_data %>% filter(Case.Control=="Control") %>% summarize(mean(shannon))

shannon_cor_data <- ggplot(Combined_data, aes(x=age_diff, y=shannon)) + geom_point() +
  facet_grid(~ Type) + geom_line(y=4.4, color="red", linetype="dashed") +
  ylab("Shannon Diversity") + xlab("Time after diagnosis")


Combined_data %>% filter(Case.Control=="Control") %>% summarize(mean(pielou_e))

even_cor_data <- ggplot(Combined_data, aes(x=age_diff, y=pielou_e)) + geom_point() +
  facet_grid(~ Type) + geom_line(y=0.636, color="red", linetype="dashed") +
  ylab("Evenness") + xlab("Time after diagnosis")


```


```{r}

sup_cor_plot <- plot_grid(faiths_cor_data, richness_cor_data, shannon_cor_data, even_cor_data,
                          nrow=2)
sup_cor_plot
```


```{r}
cor.test(Breast_data_comb$age_diff, Breast_data_comb$faith_pd, na.rm=T, method = "spearman")
cor.test(Breast_data_comb$age_diff, Breast_data_comb$shannon, na.rm=T, method="spearman")
cor.test(Breast_data_comb$age_diff, Breast_data_comb$pielou_e, na.rm=T, method="spearman")
cor.test(Breast_data_comb$age_diff, Breast_data_comb$observed_otus, na.rm=T, method="spearman")


cor.test(Prostate_data_comb$age_diff, Prostate_data_comb$faith_pd, na.rm=T, method = "spearman")
cor.test(Prostate_data_comb$age_diff, Prostate_data_comb$shannon, na.rm=T, method="spearman")
cor.test(Prostate_data_comb$age_diff, Prostate_data_comb$pielou_e, na.rm=T, method="spearman")
cor.test(Prostate_data_comb$age_diff, Prostate_data_comb$observed_otus, na.rm=T, method="spearman")

cor.test(CRC_data_comb$age_diff, CRC_data_comb$faith_pd, na.rm=T, method = "spearman")
cor.test(CRC_data_comb$age_diff, CRC_data_comb$shannon, na.rm=T, method="spearman")
cor.test(CRC_data_comb$age_diff, CRC_data_comb$pielou_e, na.rm=T, method="spearman")
cor.test(CRC_data_comb$age_diff, CRC_data_comb$observed_otus, na.rm=T, method="spearman")

```